|

### 2023-06-19 chapter_2/src/prr/color_influence_pre-reg_anon.qmd
I hate submodules

# How do colors convey political information and effect individual attitudes? {#sec-2}
<!--
    - Description: QMD script for chapter 1
    - Updated: 2022-12-28
    - Updated by: dcr
    - Notes:
        * Code:
            ** pre-test:
                *** 00_pre-test.py
                *** table called pre-test
            ** Scraping:
                *** CAPD
                    **** Execute 01_capd_text_scraping.ipynb interactively
                        ***** table called ch_1_capd_yard_signs
                    **** Execute 02_capd_img_scraping.ipynb interactively
                        ***** table called ch_1_capd_yard_signs
                    **** Execute 03_capd_downloading_images.py
                        ***** does so at end of 02_capd_img_scraping.ipynb
                    **** Execute 04_mit_election_lab_merge.py
                        ***** does so in the some-script-execution block below
        * Writing:
            ** First should write an annotated bibliography for this chapter
            ** Once annotated bibliography is complete, should have enough to start working on a pre-registered report for this chapter

-->

# Introduction

::: {#fig-pre-2000-logos layout-nrow="2"}
![Republicans pre-2000](../../assets/rnc_logo_pre-2000.png){#fig-republican}

![Democrats pre-2000](../../assets/dnc_logo_1960.png){#fig-democrat}

![Republicans 2023](../../assets/rnc_logo_2023.png){#fig-republican-post}

![Democrats 2023](../../assets/dnc_logo_2023.png){#fig-democrat-post}

Party logos
:::

Are colors important to politics? I argue that they convey information that matters to campaigns and voters. People rely on more than purely political information to understand politics [@pietraszewski_et-al_2015_c]. To reduce cognitive load, the affiliation detection system allows for efficient social behavior prediction by associating various cues with a group. In politics, associating Black voters with the Democratic party and White voters with the Republican party is one example [@pietraszewski_et-al_2015_c]. As Americans consistently link the colors red and blue with the Republican and Democratic parties [@elving_2014_npr;@williams_et-al_2022_jomp], as well as their connection to ideology [see @maestre_medero_2022_pr for example], we should explore whether color acts as an affiliative cue for individuals in American politics.

Existing theories of political information processing have yet to examine visual information more broadly. As a simple yet ubiquitous piece of visual information, color may have more influence over shaping political views than we give it credit. As visual information is processed automatically and faster than other forms of information [see @ames_et-al_2012_ohsn], I argue that it triggers associated neurological pathways that shape the subsequent information processing of the more traditionally considered types of political information. Such a process implies that the presence of politically-relevant colors has the potential to, directly and indirectly shape political attitudes.

I propose two studies to test these claims. Yard signs are a simple yet effective source of campaign branding for political candidates. They shape the attitudes of those who view them [@makse_et-al_2019_oup], and they can even aggregatively shape electoral outcomes [@green_et-al_2016_es]. As yard signs are a simple and static form of campaign advertising, they cannot rely heavily on more complicated forms of information. However, one form of information they rely on and vary in is their color. With these characteristics, yard signs are an ideal candidate both from a statistical standpoint and also substantive. From a statistical standpoint, as they are static forms of party branding they have fewer moving parts, reducing the statistical confounds that may increase Type I errors. From a substantive standpoint, while we understand that yard signs matter to campaigns, we have yet to systematically explore more simple design choices such as color.

In [Study 1](#sec-study-1), I propose an experiment that examines whether the public notice and whether their attitudes are shaped by "Republican Red" and "Democratic Blue" on the yard signs of a fictional candidate. In [Study 2](#sec-study-2), I examine whether campaigns react to the partisan composition of an electorate in recognition of color's importance to voter decision-making and attitude formation. First, I have informal interviews with campaign staff to determine what shapes the decisions about what colors to include on a campaign's yard sign.Second, I do this by examining how electoral context shapes variation in using these partisan colors.

I then present a pre-test that experimentally tests whether yard signs using "Republican Red," "Democrat Blue," and White lead to different perceptions of the candidate's political views. I find that a fictional candidate with a background of "Republican Red" is perceived to be more Republican, and a fictional candidate with a background of "Democrat Blue" are perceived to be more Democratic. I also find that Republican voters are more likely to report a willingness to vote for the candidate with "Republican Red" on the yard sign, and Democratic voters are more likely to vote for the candidate with "Democrat Blue" on the yard sign.

# The role of visual information in politics

A central theme of an edited issue of *The International Journal of Press/Politics* is that visual politics is important and understudied [@lilleker_2019_pm]. Those who are engaged in these questions attribute these challenges to methodological sophistication and the difficult task of interdisciplinary theorizing [@gerodimos_2019_pm; @bucy_joo_2021_ijpp].

The rise of television consumption changed the focus of the medium in the research for political communication scholars [@hall-jamieson_2014_ohpc]. In a similar way, so does the rise of image-based social media and has set new agendas. For several methodological and disciplinary reasons, the visual aspects of television were of little focus in the literature [@bucy_joo_2021_ijpp]. However, the news organizations and politicians are responding to the ubiquitous use of the public to use image-dominant social media platforms like TikTok and Instagram, by joining such platforms. They are relatively active on these platforms as well. Scholars need to make this transition to integrate the role of simple visual information into our theories of political information processing (abbreviated as *pip*) and attitude formation.

How does politically-relevant visual information matter to politics? From an evolutionary-biological perspective, visual information has been a common source of information for millions of years that a variety of single-and-multi-cell organisms rely upon to evaluate their environment [see @grabe_bucy_2009_oup, Chapter 1 for a useful discussion]. Visual information processing, as an ancient biological invention, the human brain is organized around the processing of it. Reflecting on this, many neuroscience scholars argue that visual information is the fastest form of information processing for humans. For example, even complex visual information, such as the warmth expressed in someone's facial features, is automatically and subconsciously processed in only about 33 milliseconds (abbreviated as ms) [@ames_et-al_2012_ohsn].

Approached from a different perspective, as humans are cognitive misers \index{cognitive misers}, visual information in the realm of politics provides efficient information to voters about politically-relevant actors and events [@lilleker_2019_pm]. Evidence suggests that voters rely on simple visual information in the background of an image to infer ideological positions [@dan_arendt_2021_ijpp] and that coverage reflects electability perceptions [@stewart_et-al_2021_ijpp] -- these influence reported desire to vote for the candidate. Images posted on social media by politicians provide more personalized information about them, and they take on their own styles [@lindholm_et-al_2021_ijpp; @peng_2021_ijpp], reflecting that it is an alternative source of information curated to attract support. Evidence suggests that even simple party branding on yard signs has an emotional appeal and encourages a shift in attitudes toward the person who owns the yard sign [@makse_et-al_2019_oup]. That is, even branding as simple as yard signs appears to influence the attitudes of those who view them. However, like all types of party branding, there remain questions about the individual visual components' contribution.

While visual politics is enjoying more attention from social scientists, there still needs to be more focus on the simplest visual information: color. In the context of the United States, the "Republican red" and "Democratic blue" is a relatively recent invention that likely has significant import in an era of significant effort by the parties to distinguish themselves from each other [@clifford_2020_pb] and where voters toe the party line [@utych_2020_es]. Since the 2000 presidential election, the media have consistently used red on their electoral maps in "horserace" journalism to represent Republicans and blue to represent Democrats [@elving_2014_npr]. The supposed consequence is that Democrats now report a preference for the color blue over the color red, and Republicans report a preference for the color red over the color blue [@schloss_palmer_2014_pbr]. Others have demonstrated the strategic choices that candidates make on branding choices -- including color -- based on the types of information they want to convey to voters to distinguish them from their rivals [see @williams_et-al_2022_jomp]. Beyond this, however, we have yet to develop and test theories about color's broader impacts on political information processing and attitude formation.

As the broader field of study of visual politics is under-theorized, theorizing about the use of color as a form of information is also quite limited -- perhaps even more so. The literature that does exist argues that colors are a source of visual information to classify more abstract concepts for voters. For example, in western Europe, voters are better at connecting the ideological positioning of a party with the color they use in their branding [@casiraghi_et-al_2022_pp]. The ability to do so is strengthened when the party is longer-surviving and is more prominent [@casiraghi_et-al_2022_pp]. The use of politically-relevant colors activates biases toward pre-existing ideological and partisan preferences among voters in a Spanish sample [@maestre_medero_2022_pr]. And male and female candidates in the United States use different colors -- among other things such as fonts -- in their campaign branding to convey distinguishing information about themselves [@williams_et-al_2022_jomp]. Despite the existing evidence, the particular psychological mechanisms driving this remain unclear. For example, it is unclear how colors may be organized into a cognitive schema that enables voters to quickly access and connect these colors to more complex political attitudes. Furthermore, it is unclear from existing evidence how colors may not just be associated with complex attitudes, but how they may enable political attitude formation and may strengthen existing connections between different political objects. As a jumping-off point, we can use the literature of political information processing (abbreviated as *pip*)\index{political information processing} to establish a common understanding of how political information is processed and how attitudes are formed by the processing of political information.

The first prominent model of *pip* derives from rational choice perspectives. The memory-based model of *pip* views attitudes as a weighted collection of prior information [see @zaller_1992_cup]. As individuals receive new information, they organize it into a schema that is relative to prior objects they already have encoded. By encoding this new information, the model predicts that individuals incorporate this new object with similar objects to form an attitude. This schema then may be more accessible in similar contexts and may then sample from its elements when prompted to express a political attitude. While the Receive-Accept-Sample (RAS) model accepts the view that expressed attitudes are a weighting of information most accessible at the time of attitude expression [@zaller_feldman_1992_ajps], it still presupposes that the weighting is an average of prior information.

The second prominent model challenges this latter point. The online model of *pip* contends that individuals do not evenly weigh information, but that whether they even store it in their long-term memory to access later is biased in the direction of supporting pre-existing attitudes [@lodge_taber_2013_cup]; this phenomenon is referred to as motivated reasoning by the psychology literature [@kunda_1990_pb]. This model suggests that people ignore new information that goes against their prior beliefs and that information confirmatory of their preferences is quicker to access -- referred to as hot cognition [@lodge_taber_2013_cup].

The online model conceptualizes this underlying information encoding and attitude retrieval mechanism as automatic [@lodge_taber_2013_cup]. This occurs due to the information's strong associations with valanced appraisals of the information guiding the attitude [@lodge_taber_2013_cup]. This brings political scientists closer to the dominant conceptualizations among neuroscientists and psychologists concerned with memory retrieval, encoding, and attitude formation [see @fazio_2007_sc]. Namely, that memories are encoded and retrieved quickly based on associations due to their association with valanced appraisals [@kensinger_fields_2022_ohhum]. What remains unexplored is how visual information such as color may prime individuals to engage in motivated reasoning pre-consciously. If they keep their attention on the object, what paths are activated by such information?

# Integrating color into a model of political information processing

Existing models of *pip* primarily focus on complex forms of political information such as text. As individuals process visual information before other sorts of information, we might expect that they may form a snap judgment or the initial appraisal of an object. This has several important implications, from suggesting that we broaden our theories of *pip* to reconsider what constitutes politically-relevant information to what it means for our substantive understanding of how people process political information and the calculus involved in political attitude formation.

Though the online information processing model goes a long way to inform us about how our physiology engages attitude formation and retrieval, the types of information it considers necessarily limits the theory's applicability to other forms of political information. Color and other simple visual information are processed much quicker and occur more frequently than text-based information [@mehta_zhu_2009_s]. As the color and other visual information are processed differently, we should also consider its use as political information. As visual information is affectively encoded [@cimbalo_et-al_1978_jgp], it can affect the affective state and processing of more complex information, such as text. The visual information provides a snap judgment or an impression of the object through faster processing and activates particular neurological processes that influence subsequent information appraisals [@ames_et-al_2012_ohsn]. This implies that the conclusions drawn from such *pip* models may be systematically biased without considering the upstream effects of non-text information; such as simple visual information like color.

Before expanding upon the role of colors in shaping political attitudes, let me first define an attitude as a concept. An attitude represents an accessible, valanced evaluation of associated prior information and experiences. This conceptualization fits with that of the Object-Evaluation Associations Model [@fazio_2007_sc]. As opposed to viewing attitudes as a latent collection of memories, as is done in the memory-based model of *pip* [@zaller_feldman_1992_ajps], it views attitudes as measurable evaluations of memories. As memories are at the core of an attitude, the association of memories with its evaluative component [see @kensinger_fields_2022_ohhum] contributes to the perspective that attitudes are affective. This implies that we should be able to measure attitudes but that such an operationalization requires careful consideration of the context's role in any given measure of an attitude as they result from memories [@fazio_2007_sc].

In line with the existing models of *pip*, I conceptualize attitudes as associative. This means that attitudes may be unstable - not stochastically, in any case. As attitudes are associative, they manifest slightly differently depending on the associative paths activated [@fazio_2007_sc]. The retrieval of relevant memories to the attitude depends on many factors, such as the recency of the event, the similarity of the context, and the importance or salience of the memory [@kahana_et-al_2022_ohhum]. This means that the memories retrieved to contribute to an attitude are quite variable. However, to understand where that variability comes from, we must understand the deeper processes influencing how information is encoded and later retrieved. This illustrates my need to build upon the reigning models of *pip*. Colors may act as a contextual feature that may lead to this variability in how a given set of political information may shape attitudes.

Colors are associative and are affectively encoded [@cimbalo_et-al_1978_jgp]. When individuals access a memory, they do not just recall an object, but they may recall visual information such as the color of an object [@mehta_zhu_2009_s]. Due to these features, colors can be processed unconsciously and very quickly [@mehta_zhu_2009_s]. As they are affectively encoded, their associations with particular memory contribute to the evaluative component of the memory. For example, individuals associate colors like red with anger and arousal [@valdez_mehrabian_1994], whereas individuals associate blue with things like happiness and pleasure [@dandrade_egan_1974_ae]. This means that colors are compelling as contextual information shaping the subsequent processing and integration of "traditional" forms of political information to construct an attitude. As the preference for a particular color correlates with political attitudes [@schloss_palmer_2014_pbr], they may have some causal influence upon political attitudes; rather than just a correlation with them.

According to the literature in affective neuroscience, visual information is processed in parts of the brain, such as the visual cortex [@goldstein_brockmole_2017_cl]. This will activate other areas of the brain and will make associated paths "hot." One such area is the amygdala. Neuroscientists believe that as visual information is quickly and subconsciously processed, the amygdala takes and appraises it based on the paths it activated; this generates a simple affective response to such information [@winkielman_et-al_2011_ohsn]. More complex, categorical emotion occurs later in conscious processing [@winkielman_et-al_2011_ohsn]. Once politically-relevant visual information is detected, the retina passes it to the brain. Once there, the brain attempts to classify the visual information by activating networks of neurons associated with the current information. With these activated pathways, the brain also attempts to appraise such information based on quick classification. For attitude formation, this implies that, as memories are affectively encoded [@kensinger_fields_2022_ohhum], these memories help areas such as the amygdala to appraise the current information.

These fast affective classifications are valanced rather than the more laborious categorization of specific emotions (e.g., anger, anxiety, fear, happy). What this means is that rather than conceptualizing the affective component of this process in line with the popular conceptualizations in political science, such as affective intelligence theory [@marcus_2000_arps], I conceptualize affect as a more extensive system that considers emotion only as a conscious classification process that occurs later than initial valanced appraisals of an object [@sander_2013_chhan; @ralph_anderson_2018_pup]. That is, emotion -- the complex classification of appraisals -- is a conscious component of affect. The pre-conscious process of affect occurs first with the simple and automatic valanced classification of an object [@winkielman_et-al_2011_ohsn; @dror_2017_er].

The pre-conscious appraisals of the visual information one encounters encourages particular behavioral and attitudinal motivations [@valentino_et-al_2011_jop; @ralph_anderson_2018_pup]. This has evolutionary roots for survival [@ralph_anderson_2018_pup; @parker_2003_p]. While affective appraisals can lead to complex motivations, such as anxiety leading to motivations for information seeking [@marcus_2000_arps], affective appraisals are valanced and are more automatic [@winkielman_et-al_2011_ohsn]. These affective appraisals lead to a desire to either retract or engage more with the object [@valentino_et-al_2011_jop]. The snap judgment resulting from the automatic processing of politically relevant visual information should likely lead to an affective response that motivates either a desire to engage more with the object or disengage.

While the visual information may encourage a particular immediate reaction to engage or disengage from the information, subsequent information processing and more conscious processing adjusts this initial appraisal generated by the snap judgment [@kensinger_fields_2022_ohhum]. While subsequent information may amend one's snap-judgment, the snap-judgment nevertheless influences the processing of subsequent information by activating particular paths, which is later encoded as associated with the object as it converts to a memory [@lodge_taber_2013_cup; @kensinger_fields_2022_ohhum].

@fig-snap-judgement-model presents an illustration of the snap-judgment model.

```{mermaid}
%%| label: fig-snap-judgement-model
%%| fig-cap: Snap judgement model
%%| fig-width: 6
%%{init: {'theme':'base', 'themeVariables':{'primaryColor':'#ffffff', 'primaryBorderColor': '#000000'}}}%%
graph TD;
    A[Detection of politically-relevant visual information] --> B(Activation of retrieval mechanism);
    B --> C[Memory and contiguous paths hot];
    C --> D[Appraisal of hot autonomous nervous system];
    D --> E1[Negative Valence];
    D --> E2[Positive Valence];
    E1 --> F1[Disengagement motivation];
    E2 --> F2[Attention activated];
    F1 --> |Neg. new information| G1[Encoded as negative];
    F1 --> |Pos. new information| G2[Encoded as positive];
    F2 --> |Neg. new information| G3[Encoded as negative];
    F2 --> |Pos. new information| G4[Encoded as positive];
    G1 --> H1[Strengthens negative path];
    G2 --> H2[Weakens negative path];
    G3 --> H3[Weakens positive path];
    G4 --> H4[Strengthens positive path];
    G2 --> |forgotten| H1;
    H2 --> |reinforced| G2;
    H3 --> |reinforced| G3;
    G3 --> |forgotten| H4;
```

All of this means for *pip* is that when we view political events or consume political information with a visual component, we are going to encode visual information along with it. Taking expectations formed from theories of motivated reasoning [see @kunda_1990_pb], I expect that the visual information that we encode with it is likely congruent with the evaluation of the object; we are likely to be unwilling to encode the visual information that is not congruent with the visual information as we do with text-based political information [see @lodge_taber_2013_cup]. That is because unconsciously, we will experience a motivation to disengage from such information as soon as we appraise the visual aspects of an object as incongruent with existing attitudes and, therefore, be uncomfortable. It should also influence how we retrieve memories when we encounter new information, which will affect the attitudes we express. Additionally, this predicts that we will spend less time processing, accessing, and encoding congruent information [@lodge_taber_2013_cup].

Let me illustrate the snap-judgment model with a common experience for residents of the United States. Say you are driving down a highway. At 65 miles an hour, you are traveling at about 95 feet per second at this speed. You split your attention. You focus your eyes on the conditions of the road in front of you, the cars in front of you, and the rear-view mirror where your kids are either dropping food in the crevice between the seats or trying to grab your attention. Out of the corner of your eye, you see a sign. It is not a road sign because it is not on the familiar white or yellow background with black lettering. It is election season. You correctly infer that it is a political yard sign. In this split second, you notice the sign's color and may see a name: Mitch McConnell. You now are racking your brain to think about who that is. If you are politically engaged, you might come to that recognition of the name quickly or it may take you significantly longer if you are less politically engaged [see @kahana_et-al_2022_ohhum]. You figure out that they are a Republican politician. You may have come to this with the help of the fact that every year you have seen yard signs on this stretch of highway. You know that when you see those electoral maps pop up on your news app on your phone that the electoral forecasts always represent Republican support with red and blue for Democrats. Once you have figured out who this person is, with the help of this other information, you react: "ugh, that guy is too loyal to Trump" or "yeah! He's loyal to Trump". You've expressed a political attitude.

What the snap-judgment model predicts is happening in your head is that as soon as the light that bounces off the sign to produce a particular wavelength hits your eyes, your brain is already trying to make sense of this information. This is a valuable tool for survival that biology has optimized for millions of years [@parker_2003_p]. Rather than slowly processing the visual information and finding yourself in the jaws of a predator or processing it quickly but forming the wrong impression and running away from a friend, the brain processes the information quickly and subconsciously [@newell_1990_hup]. To make sense of such information, it accesses familiar information similar to what it is currently attempting to process for efficiency [@kahana_et-al_2022_ohhum]. This is accessing memories and contains valanced information [@kensinger_fields_2022_ohhum]: should I avoid this, or is it pleasant? Once the brain has finished such processing, it can pass its prediction to your conscious memory. Once you form a reflex of avoid or approach, this opens up space for your brain to process the more complex information: to take the patterns of the light as shapes that construct symbols and letters. This comes later because this information not only requires access to information about what it *is* but also what it *means*; once you understand what it means, you have the information necessary to evaluate it.

The snap-judgment model predicts that you first process the colors of the yard sign. You access associative memory to figure out what those particular wavelengths represent: red, white, blue? As these colors are associated with different emotional states [see @cimbalo_et-al_1978_jgp] and the resulting behavioral consequences, your brain starts sending signals to the rest of your body to prepare it to react [see @sander_2013_chhan; @dror_2017_er]. You now need to figure out what the rest of that information was. What were the patterns of that light? It appears that there were some white letters on the sign. There was an "E," a "L," an "E," a "C," and a "T." That creates the word "ELECT." Meaning to vote for. There were some more letters on the sign: an "M," an "I," a "T," a "C," and an "H." A name. The full name is "Mitch McConnell". Since it is about politics, it must be a politician named Mitch McConnell. Now imagine the information was the same, except the color was blue. You may take more time to figure out how that Mitch McConnell person is and come to your reaction to seeing their yard sign. This is because, without the color red, you first are thinking about Democrats who are named Mitch McConnell. Only when you come up empty on your mental Rolodex do you figure out that it is the Republican Mitch McConnell.

How do you react to the color and then to the name? Social groupings are not simply abstract concepts invented by social psychologists; our neurobiology reflects them. For example, researchers find activation of parts of the brain, such as the anterior insula, when we see someone in our social group outperformed by someone from the out-group [see @zink_barter_2012_ohsn]. The anterior insula activity is associated with physical and emotional pain, not just for ourselves but also for others [@adolphs_janowski_2011_ohsn]. Others have also observed that when seeing someone part of a high-status social group, there is an increase in activity in the sensorimotor cortex and supplementary motor area, indicating more activity in the areas of the brain that encourage movement [@zink_et-al_2008_n]. Visual information about someone in your social group speeds up processing, is more salient, and demands more attention than visual information about an object outside your social group [@zink_barter_2012_ohsn].

There is significant evidence supporting the theory that our partisan identification reflects more than just our attitudes about politics but a social identity [see @campbell_et-al_1960_jws; @mason_2018_cup] that guides our attitudes [see @achen_bartels_2016_pup; @white_et-al_2014_apsr; also @bullock_2011_apsr]. As our political attitudes reflect shared views among co-partisans [@pickup_et-al_2020_pb], the congruence between political information and our partisan identification influences our reactions to such political information. This means that the visual information we glean from politics will likely motivate those neurological features of social groups and explain the resulting behavioral manifestations reacting to such information. That information is also likely to be processed at different rates as well. That is, while visual information carries general affective associations, we should also expect associations between politically-relevant colors with partisan identification.

+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+
| Hypotheses   | Expectation                                                                                                                                         |
+==============+:===================================================================================================================================================:+
| $H_1$        | People notice color                                                                                                                                 |
+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+
| $H_2$        | Colors shape perceptions of candidate based on partisan associations                                                                                |
+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+
| $H_3$        | Candidates using Red are more supported by Republicans; candidates using Blue are more supported by Democrats                                       |
+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+
| $H_4$        | Republicans spend less time evaluating candidates using Red; Democrats spend less time evaluating candidates using Blue                             |
+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+
| $H_5$        | Campaigns recognize the importance of color to voters' evaluations of candidates and respond strategically                                          |
+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+

@@

From this discussion, I expect the following: that the average potential voter pays attention to the colors used in campaign branding ($H_{1}$); that these colors that they notice shape perceptions about the person and ideological symbol represented in the branding -- meaning that they express different levels of preference for receiving more information that is similar to what they saw and their levels of preference for supporting such a campaign ($H_{2}$); that the consistency of information explains more positive perceptions -- simple visual information with more complex "traditional" information ($H_{3}$); this positive and consistent information is processed quicker than negative and inconsistent, negative and consistent, and positive and inconsistent information ($H_{4}$); and finally that campaigns make strategic choices about their branding to attract voters ($H_{5}$) in line with their primary objective of reelection [@fenno_1973_lb; @mayhew_1974_yup]. A summary of these hypotheses are included in @tbl-hypotheses.

@@

```{r}
#| label: setup-pre-test-block

    #* Set seed
set.seed(12062022)
# Source cleaning script
source("01_pre-test_cleaning.R")
    #* Load functions
box::use(
    data.table[...],
    modelsummary[modelsummary, datasummary_skim],
    cmdstanr[...],
    brms[bf, prior, brm, rename_pars, make_stancode, make_standata, pp_check, cumulative, bernoulli],
    rstan[read_stan_csv],
    marginaleffects[avg_slopes, posterior_draws, plot_slopes],
    ggplot2[ggplot, aes, labs, theme_minimal, scale_y_discrete, geom_boxplot],
    bayesplot[color_scheme_set, bayesplot_theme_set, mcmc_combo, mcmc_areas, pp_check],
    ggdist[stat_halfeye]
)
#cmdstanr::set_cmdstan_path("C:\\cmdstan")
    #* default theme
#red_custom <- c("#DCBCBC", "#C79999", "#B97C7C", "#FFFFFF", "#8F2727", "#7C0000")
color_scheme_set(scheme = "gray")
bayesplot_theme_set(new = theme_minimal())
    #* default tails
options("marginaleffects_posterior_interval" = "hdi")

@@

data[["clean"]][ # select the following columns
    , .(Female, White, age, PartyId, BlueTreatment, RedTreatment, Vote)
] |>
datasummary_skim( # make a table of descriptive statistics
    notes = c( # add some notes to the table
        "Data source: Pre-test experiment.",
        '"Unique" column includes missing values.'
    )
)
```

@@

source("04_pre-test_party_model.R")
```

@@

PartyAMES <- avg_slopes(
  PartyModel,
  type = "link"
) |>
posterior_draws()

ggplot(
  data = PartyAMES,
  aes(
    x = draw,
    y = term
  )
) +
stat_halfeye(
  slab_alpha = 0.45
) +
theme_minimal() +
labs(
  x = "Average Marginal Effect",
  y = "",
  caption = "Data source: Pre-test.\n Distribution of AME's for model posterior draws using HDI's.\n Dot reflects median posterior draw where bars reflect 90% and 95% credible intervals."
) +
scale_y_discrete(
  labels = c(
    "Blue Treatment",
    "Red Treatment"
  )
)
```

@@

```{r}
#| label: pre-test-vote-model
#| include: false

source("05_pre-test_vote_model.R")
```

```{r}
#| label: fig-pre-test-vote-cmes
#| layout-nrow: 2
#| fig-cap: "Effect of yard sign color on candidate evaluation"
#| fig-subcap:
#|  - "Republicans would vote for candidate with Red yard sign"
#|  - "No difference among partisans in support for candidate with Blue yard sign"
#| fig-width: 6
red_cme <- plot_slopes(
    VotePreTestModel,
    variables = "RedTreatment",
    condition = "PartyId",
    type = "link"
) +
theme_minimal() +
labs(
    x = "Party Identification",
    y = "CME of Red Treatment",
    caption = "Data source: Pre-test.\n Conditional Marginal Effects of Red Treatment upon support for candidate.\n Uncertainty reflected by predicted posterior draws at the 95% level with HDI's."
)

>>>>

---
title: |
  Pre-analysis plan
subtitle: |
  How do colors convey political information and affect individual attitudes?
#code-repo: |
#  For replication, go to: <https://github.com/DamonCharlesRoberts/color_as_political_information>.
author:
  - name: Anonymized for peer review
  #- name: Damon C. Roberts
  #  email: damon.roberts-1@colorado.edu
  #  orcid: 0000-0002-4360-3675
  #  title: PhD Candidate
  #  affiliations:
  #    - id: CU
  #      name: University of Colorado Boulder
  #      department: Political Science
  #      address: 333 UCB
  #      city: Boulder
  #      region: CO 
  #      postal-code: 80309-0333
abstract: |
  Colors are important to politics as a form of political information. Building upon existing theories of political information processing, attitude formation, and affect in neuroscience, I present a snap-judgement model of political information processing. In this model, colors provide automatic information about a politically-relevant object that may shape subsequent processing of more complex information. The model has important implications for how we consider the role that visual information has on political information processing and attitude formation. The model additionally provides clarity on motivations behind party branding and the ways in which information may activate partisan biases pre-consciously. I test this model using a survey experiment tracking participants' mouse movements to view different parts of yard signs that vary the use of partisan and non-partisan colors. I additionally leverage congressional redistricting to examine changes in color use for political yard signs as a strategic choice for attracting voters by communicating partisanship.
bibliography: "../../../assets/references.bib"
format:
  hikmah-pdf:
    # Spit out in drafts directory
    latex-output-dir: "../../out/prr"
    # Use biblatex-chicago
    biblatex-chicago: true
    biblio-style: authordate
    biblatexoptions: 
      - backend=biber
      - autolang=hyphen
      - isbn=false
      - uniquename=false
code-overflow: wrap
nocite: |
  @quarto
execute:
  echo: false
  warning: false
  message: false
  cache: false
params:
  replicate: false
---

+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+
| Hypotheses   | Expectation                                                                                                                                         |
+==============+:===================================================================================================================================================:+
| $H_1$        | People notice color                                                                                                                                 |
+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+
| $H_2$        | Red is associated with Republicans and Blue is associated with Democrats                                                                            |
+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+
| $H_3$        | Candidates using Red are more supported by Republicans; candidates using Blue are more supported by Democrats                                       |
+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+
| $H_4$        | Republicans spend less time evaluating candidates using Red; Democrats spend less time evaluating candidates using Blue                             |
+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+
| $H_5$        | Campaigns recognize the importance of color to voters' evaluations of candidates and respond strategically                                          |
+--------------+-----------------------------------------------------------------------------------------------------------------------------------------------------+

@@

From my review of the literature in political science, neuroscience, and communication, I expect the following: that the average potential voter pays attention to the colors used in campaign branding ($H_{1}$); that these colors that they notice shape perceptions about the person and ideological symbol represented in the branding -- meaning that they express different levels of preference for receiving more information that is similar to what they saw and their levels of preference for supporting such a campaign ($H_{2}$); that the consistency of information explains more positive perceptions -- simple visual information with more complex "traditional" information ($H_{3}$); this positive and consistent information is processed quicker than negative and inconsistent, negative and consistent, and positive and inconsistent information ($H_{4}$); and finally that campaigns make strategic choices about their branding to attract voters ($H_{5}$) in line with their primary objective of reelection [@fenno_1973_lb; @mayhew_1974_yup]. A summary of these hypotheses are included in @tbl-hypotheses.

@@

```{r}
#| label: setup-pre-test-block
# set working directory
#setwd("./chapter_2/src/prr")
# Set seed
set.seed(12102022)
#setwd("./chapter_2/src/prr")
# Source cleaning script
source("eda/pre-test/01_pre-test_cleaning.R")
# Load relevant functions
if (params$replicate == TRUE){
  box::use(
    modelsummary[
      datasummary_skim
    ]
    ,marginaleffects[
      avg_slopes
      ,posterior_draws
      ,plot_slopes
    ]
    ,ggplot2[
      ggplot
      ,aes
      ,theme_minimal
      ,labs
      ,scale_y_discrete
    ]
    ,ggdist[
      stat_halfeye
    ]
    ,brms[
      brmsformula
      ,cumulative
      ,prior
    ]
    ,rstan[stan_model]
    ,./R/sim_all[sim_all]
    ,./R/discrepancy_df[discrepancy_df]
    ,./R/true_positive_hist[true_positive_hist]
  )
} else {
  box::use(
    modelsummary[
      datasummary_skim
    ]
    ,marginaleffects[
      avg_slopes
      ,posterior_draws
      ,plot_slopes
    ]
    ,ggplot2[
      ggplot
      ,aes
      ,theme_minimal
      ,labs
      ,scale_y_discrete
    ]
    ,ggdist[
      stat_halfeye
    ]
    ,./R/discrepancy_df[discrepancy_df]
    ,./R/true_positive_hist[true_positive_hist]
  )
}
  #* default tails
options("marginaleffects_posterior_interval" = "hdi")

@@

df_select <- data[["clean"]][ # select the following columns
  , c("Female", "White", "age", "PartyId", "BlueTreatment", "RedTreatment", "Vote"), with = FALSE
] |> as.data.frame()

datasummary_skim( # make a table of descriptive statistics
  df_select
  , notes = c( # add some notes to the table
      "Data source: Pre-test experiment.",
      '"Unique" column includes missing values.'
  )
  , histogram = FALSE
  , output = "latex"
)
```

@@

if (params$replicate == TRUE) {
  PartyFormula <- bf(
    Party ~ RedTreatment + BlueTreatment
    ,family = cumulative(link = "logit")
  )
  PartyModel <- brm(
    formula = PartyFormula,
    ,data = data[["clean"]]
    ,chains = 1
    ,silent = 0
  )
  # Store model object
  save(
    PartyModel
    ,file="../../data/models/prr/pre-test/pre-test_party_model.RDS"
  )
} else (
  # load the stored model object
  load(
    file="../../data/models/prr/pre-test/pre-test_party_model.RDS"
  )
)

@@

partyAMES <- avg_slopes(
  PartyModel,
  type = "link"
) |>
posterior_draws()

ggplot(
  data = partyAMES,
  aes(
    x = draw,
    y = term
  )
) +
stat_halfeye(
  slab_alpha = 0.45
) +
theme_minimal() +
labs(
  x = "Average Marginal Effect",
  y = "",
  caption = "Data source: Pre-test.\n Distribution of AME's for model posterior draws using HDI's.\n Interpreted as the logged odds of labelling the candidate as a partisan\n given whether they viewed a red yard sign or blue yard sign\n while holding all other variables at their mean value.\n Dot reflects median posterior draw where bars reflect 90% and 95% credible intervals."
) +
scale_y_discrete(
  labels = c(
    "Blue Treatment",
    "Red Treatment"
  )
)
```

@@

```{r}
#| label: pre-test-vote-model
#| include: false
if (params$replicate == TRUE) {
  # fit the model
  VotePreTestFormula <- bf(
    Vote ~ RedTreatment + BlueTreatment + PartyId + RedTreatment * PartyId + BlueTreatment * PartyId
    ,family = cumulative(link = "logit")
  )
  VotePreTestModel <- brm(
    formula = VotePreTestFormula
    ,data = data[["clean"]]
    ,chains = 1
    ,silent = 0
  )
  # Store model object
  save(
    VotePreTestModel
    ,file="../../data/models/prr/pre-test/pre-test_vote_model.RDS"
  )
} else {
  # load the stored model object
  load(
    file="../../data/models/prr/pre-test/pre-test_vote_model.RDS"
  )
}
```
